---
title: "p8105_hw1_hl4104"
date: "2026-09-19"
author: "Houson Lin" 
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)

#data import
data("penguins", package = "palmerpenguins")
```

## Problem 1
```{r}
glimpse(penguins)
```
By the `glimpse()` function above, we could tell that:

The `penguins` dataset contains measurements for individual penguins observed on islands in the Palmer Archipelago and has **`r nrow(penguins)` rows** and **`r ncol(penguins)` columns**. Each row represents one penguin, and each column represents one characteristic.

The main variables are:
- `species`: penguin species (`r paste(levels(penguins$species), collapse = ", ")`);
- `island`: island where the penguin was observed (`r paste(levels(penguins$island), collapse = ", ")`);
- `bill_length_mm` and `bill_depth_mm`: bill measurements in millimeters;
- `flipper_length_mm`: flipper length in millimeters;
- `body_mass_g`: body mass in grams;
- `sex`: recorded sex (`r paste(levels(penguins$sex), collapse = ", ")`);
- `year`: year of observation, from `r min(penguins$year, na.rm = TRUE)` to `r max(penguins$year, na.rm = TRUE)`.

There are `r sum(is.na(penguins$flipper_length_mm))` missing values in `flipper_length_mm`. After excluding them with `na.rm = TRUE`, the mean flipper length is **`r round(mean(penguins$flipper_length_mm, na.rm = TRUE), 2)` mm**.

```{r penguin-scatter}
penguin_plot <- ggplot(
  penguins,
  aes(x = bill_length_mm, y = flipper_length_mm, color = species)
) +
  geom_point(alpha = 0.35, size = 2, na.rm = TRUE) +
  theme_minimal(base_size = 12)

penguin_plot

#plot saving
ggsave(
  filename = "penguins_scatterplot.png",
  plot = penguin_plot,
  width = 8,
  height = 5,
  units = "in",
  dpi = 300
)
```
## Problem 2
```{r creating data with different types}
set.seed(8105) # ensuring the unique creation

created_data <- tibble(
  normal_sample = rnorm(10),
  above_zero = normal_sample > 0,
  character_value = as.character(1:10),
  category = factor(
    c("l", "m", "h", "l", "m", "h", "l", "m", "h", "l"),
    levels = c("l", "m", "h") #here l for low, m for medium, h for high
  )
)

created_data
```

```{r mean calculation}
mean(created_data %>% pull(normal_sample))
mean(created_data %>% pull(above_zero))
mean(created_data %>% pull(character_value))
mean(created_data %>% pull(category))
```
Upper two lines of code work and the rest do not, giving warnings that 'argument is not numeric or logical: returning NA'.


```{r explicit-coercion, results='hide'}
as.numeric(created_data %>% pull(above_zero))
as.numeric(created_data %>% pull(character_value))
as.numeric(created_data %>% pull(category))

# try mean again
mean(as.numeric(created_data %>% pull(character_value)))
mean(as.numeric(created_data %>% pull(category)))
```
'as.numeric' function modify the type of those columns into numeric, which gives 0 or 1 to boolean type, numeric value to characters, 123 to different factor levels, and thus they could be calculated by 'mean' and it works.
